
Topic:  NYC Energy Model
==========================
Author: Shaky Sherpa
Date:   Decomber 1 2013
Course: GTECH 705 - Spatial Data Analysis
Final Project 
----------------------------------------------
Here's the general steps i am performing:
1. Define neighbors (k nearest w/point data)
2. Create weights matrix
3. Moran's test of observed, Moran scatterplot
4. Run OLS regression
5. Check residuals for spatial dependence
6. Determine which SR model to use w/LM tests
7. Run spatial regression model


```{r}
# Clear the workspace
rm(list=ls())
#load required libraries
library(spdep)
library(spgwr)
library(rgdal)
library(maptools)
library(ggplot2)
```

# Spatial regression workflow
```{r}
# Load some data
zip.shp = readOGR("data/zip_energy.shp","zip_energy")
# reproject the data to NAD83 / New York Long Island (ftUS)
zip <- spTransform(zip.shp, CRS("+init=epsg:2263"))

# Remove missing  data and keep only complete records
zip.data = zip[complete.cases(zip@data),]



# Check what this does
IDs = row.names(zip.data)
```


#Regression model
```
energy = elec_kwh ~ pop + medianinco + avghomeval + houseunits
```
#Create weight matrices
```{r}
# Create neighbors list
zip.data.gal.nb = poly2nb(zip.data,snap=5280 )

# Create weights matrix
zip.data.listw = nb2listw(zip.data.gal.nb, style="W")
```
#plotting

```{r fig.width=7, fig.height=6}
par(mar=c(0,0,0,0))
plot(zip.data)
plot(zip.data.listw, coords=coordinates(zip.data), add=TRUE)
```

# Test for spatial autocorrelation

```{r}
# Moran's Test
moran.test(zip.data$elec_kwh, listw=zip.data.listw, alternative="two.sided", zero.policy=TRUE)
```

# Test for local spatial autocorrelation

```{r fig.width=7, fig.height=6}
moran.plot(zip.data$elec_kwh, zip.data.listw, ylab="Spatially lagged Energy Consumption", xlab="Energy consumption in each zip")
```

```{r}
zip.data.li = localmoran(zip.data$elec_kwh, zip.data.listw)
zip.data$localm = zip.data.li[,4]

```

# Plot
```{r fig.width=7, fig.height=6}
spplot(zip.data, "localm", main="Local Moran's Ii Z-Scores")
```

# Run OLS Regression

```{r}
zip.data.lm = lm(elec_kwh ~ pop + medianinco + avghomeval + houseunits, data=zip.data)
summary(zip.data.lm)
```
# Check residuals for spatial dependence

```{r}
zip.data$lmres = residuals(zip.data.lm) # Grab the residuals (though don't really need to do this)
lm.morantest(zip.data.lm, zip.data.listw)
```
# Perform lagrage multiplier test
# Robust tests used to find a proper alternative
# Only use robust forms when BOTH LMErr and LMLag are significant

```{r}
lm.LMtests(zip.data.lm, zip.data.listw, test="all")
```

```{r}
install.packages("lmtest")
library(lmtest)
bptest(zip.data.lm)
```

# Indicates errors are heteroskedastic
# Not surprising since we have spatial dependence!
# Fit spatial regression models

# Spatial lag model(Spatial Simultaneous Autoregressive lag model estimation)

```{r}
zip.data.lag = lagsarlm(elec_kwh ~ pop + medianinco + avghomeval + houseunits, data=zip.data, listw=zip.data.listw)
summary(zip.data.lag)
```


# Some more diagnostics
```{r}
bptest.sarlm(zip.data.lag)#Breusch-Pagan test
```

# LM test suggests there is no more spatial autocorrelation in the data
# BP test indicates remaining heteroskedasticity in the residuals
# Most likely due to misspecification
# Spatial error model(Maximum likelihood estimation of spatial simultaneous autoregressive error models of the form)

```{r}
zip.data.err = errorsarlm(elec_kwh ~ pop + medianinco + avghomeval + houseunits, data=zip.data, listw=zip.data.listw)
summary(zip.data.err)
bptest.sarlm(zip.data.err)
```
#Pick the model that has lower AIC value
#
## GWR Model (Instead of fitting one model we account for fitting model with multiple variables)
##
```{r}
zip.data.lm = lm(elec_kwh ~ pop + medianinco + avghomeval + houseunits, data=zip.data)
zip.data$lmres = residuals(zip.data.lm) # Extract Model Residuals
```

#Crossvalidation of bandwidth for GWR
```{r}
#select the optimized bandwidth
zip.data.bw = gwr.sel(elec_kwh ~ pop + medianinco + avghomeval + houseunits, data=zip.data)
zip.data.gauss = gwr(elec_kwh ~ pop + medianinco + avghomeval + houseunits, data=zip.data, bandwidth=zip.data.bw, hatmatrix=TRUE)
summary(zip.data.gauss)
```
#Plotting the variables
```{r}
spplot(zip.data, "elec_kwh", main="Energy consumption in 2010")
spplot(zip.data, "energy_per", main="Energy use per person")
spplot(zip.data, "energy_hh", main="Energy use per household")
spplot(zip.data, "pop_densit", main="Population Density in each zipcode")
spplot(zip.data, "medianinco", main="Median Income in each zipcode")
spplot(zip.data, "houseunits", main="Houseunits  in each zipcode")
spplot(zip.data, "avghomeval", main="Average home value in each zipcode")
```

#Plot the LM Residuals
```{r fig.width=7, fig.height=6}
spplot(zip.data, "lmres", main="Linear Model Residuals")
```
#Plot Gaussian 
```{r fig.width=7, fig.height=6}
spplot(zip.data.gauss$SDF, "energy_per", main="Energy use per person")
```
```{r fig.width=7, fig.height=6}
spplot(zip.data.gauss$SDF, "energy_hh", main="Energy use per household")
```
```{r}
res = zip.data.gauss$SDF
zip.data$gwrres = zip.data$elec_kwh - res$pred
```
```{r fig.width=7, fig.height=6}
spplot(zip.data, "gwrres", main="GWR Model Residuals")
```

